Developmental trajectories of word recognition: a priming study in bilinguals

### 2023-11-05 manuscript/manuscript.qmd
General discussion

One of the most interesting pieces of evidence in support of a cascaded account of lexical processing is provided by bilinguals. A critical property the bilingual lexicon is that its representations are accessed in a language-non selective way. Even during monolingual situations, bilinguals access their lexical representations in both of their languages [e.g., @thierry2007brain; @dunabeitia2009masked]. This is reflected in bilinguals' performance in word recognition and production tasks. For instance, @marian1999activation presented Russian-English bilinguals with a series of trials that started with an auditory instruction like "Poloji marku nije krestika" ["Put the stamp below the cross"]. The instruction would be present in Russian (dominant language for most participants). The target object (*marku* [*stamp*], in this case) would be present in the screen, together with an object whose English label shared phonological onset with the target label (e.g., *mark*), and two unrelated distractors whose labels in Russian and English did not shared phonological relationship with the target label. Converging with @allopenna1998tracking's results in monolingual adults, after hearing the target label bilinguals fixated the cross-language phonological distractor significantly more than the unrelated distractors. These findings suggest that participants activated phonologically related word-forms in both Russian *and* English, which affected their overt visual exploration patterns during the task.

Language non-selectivity in the bilingual lexicon has also been shown to affect top-down lexical access during word production. @costa2000cognate presented Catalan-Spanish bilinguals with a series of pictures of familiar objects. For each object, participants were asked to name it in Catalan (their dominant language). Unbeknownst to participants, the authors manipulated the phonological overlap between the pictures' labels in Catalan and their translations in Spanish. Participants produced faster words that shared phonological overlap with their Spanish translations (e.g., *gat*-*gat* [*cat*]) than those unrelated to their translation (e.g., *taula*-*mesa* [*table*]). Critically, Spanish monolinguals who completed the same task, naming the pictures in Spanish, showed equivalent naming times in both conditions. These results reveal that bilinguals activated their Spanish phonology, despite performing the naming task exclusively in Catalan: the visual recognition of the presented pictures led to the activation of its associated phonological forms in both language sin parallel, which influenced the subsequent the dynamics of word production.

@@

Using a similar task, @floccia2020translation provided conflicting results to those of @von2012language. The authors tested 27-month-old simultaneous bilinguals in a cross-language priming paradigm. The two languages participants were learning were English, and an Additional Language (either Cantonese, Dutch, French, German, Greek, Italian, Mandarin, Polish, Portuguese, or Spanish). A prime word was embedded at the end of a carrier sentence that participants listened to. Then, participants were presented with the auditory label of the target word, and two pictures were shown side-by-side, the target picture, and a distractor picture. In Study 1, the target word was the translation equivalent of the prime word (*cheese*-*fromage* for an English-French bilingual). In other trials, prime and target words were unrelated (e.g., *sock*-*fromage*). There was no phonological overlap at onset between prime and target or distractor labels in either of the two conditions. Participants were presented with two blocks of related and unrelated trials. In one of the blocks, participants were presented with English prime words and Home Language target words (*sock*-*fromage*), and in the other block they were presented with Additional Language prime words and English target words (*fromage*-*sock*). The authors found a cross-linguistic priming effect, in which participants showed a stronger preference for the target picture in the post-naming phase in related trials, as compared to unrelated trials. This effect was found regardless of the language of the prime (English or Home Language), or the dominance of the language of the prime (i.e., whether the prime word belonged to the dominant language). Target preference exceeded chance level in both conditions. In Study 2, the authors tested a similar group of bilinguals in a cross-linguistic semantic priming task, in which participants were primed with words in English, and tested with semantically related words in the Additional Language (or vice versa). This time, prime and target words were not translation equivalents, but rather words whose referents shared semantic features. Results from Study  2 were parallel to those from Study 1. The authors found a strong semantic priming effect across languages, regardless of the language of the prime or the dominance of the language of the prime. Overall these results suggest that bilinguals recognised target words in both conditions, and benefited from a cross-linguistic facilitation effect from translation equivalents. The symmetry of such effect across both languages suggests that participants lexical access occurred in parallel for the two members of translation equivalents. As @floccia2020translation note, the stimuli set of Study 2 contains multiple cognates, especially in languages with a larger degree of phonological overlap across translation pairs, like English and Dutch. The authors run *post hoc* analyses exploring the effect of cognateness on the cross-language priming effects found, which showed no evidence of a cognateness effect on target preference. However, the cross-language effects reported in Study 2 might still reflect the confounding influence of cognateness, as the resulting effect might have been generated by an interaction by the phonological and semantic overlap between cognate primes and targets.

@@

Priming-though-translation paradigms, however, come with a cost. When prime words are auditorily presented, the task constrains the naming context to one of the languages. The consequence is a potential asymmetry between the activation dynamics of the lexical representation of both labels: while the participant has heard one label, the activation of its translation must spread across semantic links. This might lead to a lower strength, and higher latency of the activation of the translation.

In the present study, we use an adaptation of an implicit naming paradigm by @mani2010infant, in which participants are primed with a picture, instead of an auditorily presented label. This paradigm prevents infants from activating the prime label in any of their two languages, and therefore makes it more likely that they will activate the prime labels in both languages in parallel. Infants implicitly named prime pictures despite such pictures having been presented in silence, and that the phonology of the resulting label interacted with the subsequent auditory recognition of the target label. In particular, the authors found a facilitation effect: target looking preference was stronger in phonologically related primes, indicating that the activated phonological segments of the generated prime label facilitated the subsequent activation of the phonological representation of the target label during recognition.

When extended to the bilingual case, this implicit naming task may offer bilingual participants with an unconstrained naming context in which the activation of both labels may occur in a truly cascaded fashion for both languages in parallel: the recognition of the prime object depicted activates the (common) semantic representation for both members of the translation equivalent, and this in turn activates the two available phonological word-forms, one in each language.

This study also provides a more suitable design for drawing conclusions about developmental change in word recognition trajectories for two reasons. First, it extends the age range of participants, compared to previous studies. Second, participants were tested at three selected age points to ensure a more balanced sample size across ages. Third, the partially longitudinal design allows modelling the performance of participants across testing sessions altogether.

In order to circumvent the problem of limited vocabulary knowledge
in the non-dominant language, we propose to test participants only in their dominant language, therefore
directing our research to the question of how possessing a second language affects processing of the
dominant one in bilingual toddlers [@costa2014does provide a fuller description of the
advantages of such approach].

The present study is aimed at investigating the developmental trajectories of word recognition in bilingualism, and how (if at all) language non-selectivity shapes them. Using an adaptation of [@mani2010infant], we presented 20- to 32-months-old toddlers with a cross-linguistic phonological priming paradigm. We manipulated both the phonological overlap between the prime word and target word, and the cognate status of the prime word, to test the hypothesis that bilingual toddlers implicitly name familiar pictures in both languages. If so, phonological priming effect should be larger when the prime label in  both languages shares phonological overlap with the target words (cognate condition), compared to when only the label of the prime in one language overlaps phonologically with the target (non-cognate condition). We also also included a control *Unrelated* condition in which regardless of the cognate status of the prime word, neither of the prime labels in any language overlapped phonologically with the target word.

We compared the explanatory power of three measures of vocabulary size on participants' trajectories of word recognition: L1 vocabulary and total vocabulary.Previous studies suggest that same-language vocabulary size predicts better bilingual infants' performance in word recognition tasks, as compared to cross-language vocabulary size [@marchman2010vocabulary]

Hypothesis 1: if lexical representations linked through inhibitory links (compete for selection), like in the adult lexicon, target looking should increase faster in *Unrelated* trials compared to *Related* trials. If lexical representations are linked through excitatory links, target preference should increase faster in *Related* trials, compared to *Unrelated* trials. If lexical representations are sparsely connected, participants should show equivalent patterns of target looking in *Related* and *Unrelated* trials.

Hypothesis 2: if bilinguals' lexical representations linked across languages through inhibitory links (compete for selection), target looking should increase faster in this group in *Related/Non-cognate* trials compared to *Related/Cognate* trials. If lexical representations are linked across languages through excitatory links, target preference should increase faster in *Related/Cognate* trials, compared to *Unrelated/Cognate* trials. If lexical representations are sparsely connected across languages, bilinguals' patterns of target looking should be equivalent for *Related/Cognate* and *Related/Non-cognate* trials. This pattern is also expected for monolingual infants.

Hypothesis 3: if the emergence of within-language or cross-language phonological links is explained (if at all) by the growth of within-language vocabulary size, the difference in target looking between *Related* and *Unrelated*, or between *Related/Cognate* and *Related/Non-cognate* trials should be larger in participants with larger vocabulary sizes in their dominant language, in which they were tested. This should be reflected by a $\text{Condition} \times \text{L1 vocabulary}$ two -way interaction, or a $\text{Condition} \times \text{Group} \times \text{Dominant vocabulary}$ three-way interaction.

Hypothesis 4: if the emergence of within-language or cross-language phonological links is explained better by the growth of cross-language vocabulary size, compared to within-language vocabulary size, a model that incorporates total vocabulary size as a predictor instead of dominant vocabulary size should show better predictive performance. If the emergence of such links is explained at all by the growth of vocabulary size, both models incorporating vocabulary size as a predictor should show better predictive performance than a model incorporating participants' age.

@@

```{r results-0}
var_labels <- c("b_Intercept" = "Intercept",
				"b_age_std" = "Age (+1 SD)",
				"b_condition1" = "Condition (U vs. RN)",
				"b_condition1" = "Condition (RN vs. RC)",
				"b_lp1" = "Group",
				"sd_child_id__Intercept" = "Intercept",
				"sd_child_id__age_std" = "Age",
				"sd_child_id__lpBilingual" = "Group",
				"sd_child_id:session_id__Intercept" = "Intercept",
				"sd_child_id:session_id__age_std" = "Age",
				"sd_child_id:session_id__lpBilingual" = "Group")

@@

```{r tbl-loos}
loos_tbl[1:5] |> 
	mutate(model = case_when(model=="fit_0" ~ "$\\mathcal{M}_{0}$",
							 model=="fit_1" ~ "$\\mathcal{M}_{1}$",
							 model=="fit_2" ~ "$\\mathcal{M}_{2}$",
							 model=="fit_3" ~ "$\\mathcal{M}_{3}$",
							 model=="fit_4" ~ "$\\mathcal{M}_{4}$",
							 model=="fit_5" ~ "$\\mathcal{M}_{5}$")) |>
	gt() |> 
	fmt_number(is.numeric) |> 
	# fmt_markdown(is.character, ) |> 
	cols_label(elpd_diff = "ELPD (diff.)",
			   se_diff = "SE ELPD (diff)",
			   elpd_loo = "ELPD",
			   se_elpd_loo = "SE ELPD")
```

@@

The outcomes of $\mathcal{M}_{0}$, which only included *Age* and as main effect, revealed successful word recognition across participants. Overall, participants' looking time exceeded chance levels, as indicated by the fact that the 95% HDI of the intercept term excluded zero ($\beta$ = `r coefs$fit_0$b_intercept$mdn_label`, 95% HDI = `r coefs$fit_0$b_intercept$ci_label`). The variability of participants' target preference varied significantly more between participants ($\sigma$ = `r coefs$fit_0$sd_child_id_intercept$mdn_label`, 95% HDI = `r coefs$fit_0$sd_child_id_intercept$ci_label`) than within testing sessions from the same participant ($\sigma$ = `r coefs$fit_0$sd_child_id_session_id_intercept$mdn_label`, 95% HDI = `r coefs$fit_0$sd_child_id_session_id_intercept$ci_label`). The 95% HDI of the coefficient of *Age* had a positive sign, but did not exclude zero ($\beta$ = `r coefs$fit_0$b_age_std$mdn_label`, 95% HDI = `r coefs$fit_0$b_age_std$ci_label`), indicating that participants from all ages showed equivalent overall target word recognition.

The 95% HDI of the coefficient of *Group* in $\mathcal{M}_{1}$ included zero ($\beta$ = `r coefs$fit_1$b_lp1$mdn_label`, 95% HDI = `r coefs$fit_1$b_lp1$ci_label`), indicating an equivalent overall target preference in monolinguals and bilinguals. The 95% HDI of the first contrast of this predictor, comparing *Unrelated* and *Related/Non-cognate* trials, included zero ($\beta$ = `r coefs$fit_2$b_condition1$mdn_label`, 95% HDI = `r coefs$fit_2$b_condition1$ci_label`). The 95% HDI of the second contrast, comparing *Related/Non-cognate* and *Related/Cognate* trials, also included zero ($\beta$ = `r coefs$fit_2$b_condition2$mdn_label`, 95% HDI = `r coefs$fit_2$b_condition2$ci_label`). The overall target preference was equivalent across both pairwise condition comparisons. The interaction term between the first *Condition* contrast contained zero ($\beta$ = `r coefs$fit_3$b_condition1_lp1$mdn_label`, 95% HDI = `r coefs$fit_3$b_condition1_lp1$ci_label`). The interaction term between the second *Condition* contrast also contained zero ($\beta$ = `r coefs$fit_3$b_condition2_lp1$mdn_label`, 95% HDI = `r coefs$fit_3$b_condition2_lp1$ci_label`). The outcomes of this model indicate that lack of differences in overall target looking time across conditions remained across monolinguals and bilinguals.

@@

@@

@@

```{r tbl-loos-oxf}
loos_tbl |> 
	mutate(model = case_when(model=="fit_oxf_0" ~ "$\\mathcal{M}_{0}$",
							 model=="fit_oxf_1" ~ "$\\mathcal{M}_{1}$",
							 model=="fit_oxf_2" ~ "$\\mathcal{M}_{2}$",
							 model=="fit_oxf_3" ~ "$\\mathcal{M}_{3}$")) |>
	gt() |> 
	fmt_number(is.numeric) |> 
	# fmt_markdown(is.character, ) |> 
	cols_label(elpd_diff = "ELPD (diff.)",
			   se_diff = "SE ELPD (diff)",
			   elpd_loo = "ELPD",
			   se_elpd_loo = "SE ELPD")
```

The outcomes of $\mathcal{M}_{0}$, which only included *Age* and as main effect, revealed successful word recognition across participants. Overall, participants' looking time exceeded chance levels, as indicated by the fact that the 95% HDI of the intercept term excluded zero ($\beta$ = `r coefs$fit_0$b_intercept$mdn_label`, 95% HDI = `r coefs$fit_0$b_intercept$ci_label`). The variability of participants' target preference varied significantly more between participants ($\sigma$ = `r coefs$fit_0$sd_child_id_intercept$mdn_label`, 95% HDI = `r coefs$fit_0$sd_child_id_intercept$ci_label`) than within testing sessions from the same participant ($\sigma$ = `r coefs$fit_0$sd_child_id_session_id_intercept$mdn_label`, 95% HDI = `r coefs$fit_0$sd_child_id_session_id_intercept$ci_label`). The 95% HDI of the coefficient of *Age* had a positive sign, but did not exclude zero ($\beta$ = `r coefs$fit_0$b_age_std$mdn_label`, 95% HDI = `r coefs$fit_0$b_age_std$ci_label`), indicating that participants from all ages showed equivalent overall target word recognition.

>>>>

One of the most interesting pieces of evidence in support of a cascaded account of lexical processing is provided by bilinguals. A critical property the bilingual lexicon is that its representations are accessed in a language-non selective way. Even during monolingual situations, bilinguals access their lexical representations in both of their languages [e.g., @thierry2007brain; @dunabeitia2009masked; @HOSHINO; @STROJKERS2010; @dijktra2010baayen]. This is reflected in bilinguals' performance in word recognition and production tasks. For instance, @marian1999activation presented Russian-English bilinguals with a series of trials that started with an auditory instruction like "Poloji marku nije krestika" ["Put the stamp below the cross"]. The instruction would be present in Russian (dominant language for most participants). The target object (*marku* [*stamp*], in this case) would be present in the screen, together with an object whose English label shared phonological onset with the target label (e.g., *mark*), and two unrelated distractors whose labels in Russian and English did not shared phonological relationship with the target label. Converging with @allopenna1998tracking's results in monolingual adults, after hearing the target label bilinguals fixated the cross-language phonological distractor significantly more than the unrelated distractors. These findings suggest that participants activated phonologically related word-forms in both Russian *and* English, which affected their overt visual exploration patterns during the task.

Language non-selectivity in the bilingual lexicon has also been shown to affect top-down lexical access during word production. @costa2000cognate presented Catalan-Spanish bilinguals with a series of pictures of familiar objects. For each object, participants were asked to name it in Catalan (their dominant language). Unbeknownst to participants, the authors manipulated the phonological overlap between the pictures' labels in Catalan and their translations in Spanish. In half of the trials, the labels associated to pictures were cognates, that is, phonologically similar translation equivalents in Catalan and Spanish, whereas in the other half of the trials the labels were non-cognates. Participants produced faster words in cognate trials (e.g., *gat*-*gat* [*cat*]) than in non-cognate trials (e.g., *taula*-*mesa* [*table*]). Critically, Spanish monolinguals who completed the same task, naming the pictures in Spanish, showed equivalent naming times in both conditions. These results reveal that bilinguals activated their Spanish phonology, despite performing the naming task exclusively in Catalan: the visual recognition of the presented pictures led to the activation of its associated phonological forms in both language sin parallel, which influenced the subsequent the dynamics of word production. the differential processing of cognates has generated a large body of literature providing support for the language-non selective account of lexical access [e.g., @degrootandnas1992; @schwartzkroll2007; @dijkstragrainger1999; @dufourkroll1995; @lemhofferetal2008]

@@

Using a similar task, @floccia2020translation provided conflicting results to those of @von2012language. The authors tested 27-month-old simultaneous bilinguals in a cross-language priming paradigm. The two languages participants were learning were English, and an Additional Language (either Cantonese, Dutch, French, German, Greek, Italian, Mandarin, Polish, Portuguese, or Spanish). A prime word was embedded at the end of a carrier sentence that participants listened to. Then, participants were presented with the auditory label of the target word, and two pictures were shown side-by-side, the target picture, and a distractor picture. In Study 1, the target word was the translation equivalent of the prime word (*cheese*-*fromage* for an English-French bilingual). In other trials, prime and target words were unrelated (e.g., *sock*-*fromage*). There was no phonological overlap at onset between prime and target or distractor labels in either of the two conditions. Participants were presented with two blocks of related and unrelated trials. In one of the blocks, participants were presented with English prime words and Home Language target words (*sock*-*fromage*), and in the other block they were presented with Additional Language prime words and English target words (*fromage*-*sock*). The authors found a cross-linguistic priming effect, in which participants showed a stronger preference for the target picture in the post-naming phase in related trials, as compared to unrelated trials. This effect was found regardless of the language of the prime (English or Home Language), or the dominance of the language of the prime (i.e., whether the prime word belonged to the dominant language). Target preference exceeded chance level in both conditions.

In Study 2, the authors tested a similar group of bilinguals in a cross-linguistic semantic priming task, in which participants were primed with words in English, and tested with semantically related words in the Additional Language (or vice versa). This time, prime and target words were not translation equivalents, but rather words whose referents shared semantic features. Results from Study  2 were parallel to those from Study 1. The authors found a strong semantic priming effect across languages, regardless of the language of the prime or the dominance of the language of the prime. Overall these results suggest that bilinguals recognised target words in both conditions, and benefited from a cross-linguistic facilitation effect from translation equivalents. The symmetry of such effect across both languages suggests that participants lexical access occurred in parallel for the two members of translation equivalents. As @floccia2020translation note, the stimuli set of Study 2 contains multiple cognates, especially in languages with a larger degree of phonological overlap across translation pairs, like English and Dutch. The authors run *post hoc* analyses exploring the effect of cognateness on the cross-language priming effects found, which showed no evidence of a cognateness effect on target preference. However, the cross-language effects reported in Study 2 might still reflect the confounding influence of cognateness, as the resulting effect might have been generated by an interaction by the phonological and semantic overlap between cognate primes and targets.

@@

Although priming-through-translation studies suggest parallel activation in participants' lexicon in both languages, their design prevents strong interpretations about the extent to which parallel activation is the result of the experimental manipulations of interest. The presentation of primes and targets in both languages in the same task---or even within the same trial--- presents participants with a bilingual context that might lead to a higher baseline activation of lexical representations of both languages, compared to monolinguals situations in which parallel activation is claimed to spontaneously take place [@GROSJEAN1997]. Another limitation of priming-through-translation studies is that, when prime words are auditorily presented, the task constrains the naming context to one of the languages. The consequence is a potential asymmetry between the activation dynamics of the lexical representation of both labels: while the participant has heard one label, the activation of its translation must spread across semantic links. This might lead to a lower strength, and higher latency of the activation of the translation.

In the present study, we aim to put the hypothesis that bilingual infants activate both languages in parallel to a harder test. We use an extension of the implicit naming paradigm by @mani2010infant, in which participants are primed with a picture presented in silence. This provides participants with an exclusively monolingual context in which they perform the task in only one of their languages. If differences in task performance are found associated with covert manipulation of cross-linguistic properties of the stimuli, such differences may be more conclusively linked to the parallel activation of lexical representations in participants' lexicons.

Another advantage of the implicit naming paradigm in the investigation of bilingual lexical processing is the fact that, since participants are not presented with any auditory prime label, the activation of the lexical representation of the prime may occur in both languages as the result of lexical retrieval after visual recognition of the prime picture. This contrasts with priming through-translation tasks, in which the prime label in one of the languages is directly activated by auditory word recognition, while the label of its translation is expected to be activated as the result of cascaded activation spreading through their semantic association. This may lead to an asymmetry in the time course of the activation of both labels, which might in turn affect the dynamics of phonological priming in subsequent word recognition.

Finally, testing bilingual populations at early ages presents the limitation that the vocabulary size of participants' non-dominant language is relatively small, which substantially constrains the number of items that can be used to as stimuli in the task, as well as the number of trials that can be constructed to generate stimuli lists. In order to circumvent the problem of limited vocabulary knowledge in the non-dominant language, tested participants only in their dominant language, therefore directing our research to the question of how possessing a second language affects processing of the dominant one in bilingual toddlers [@costa2014does provide a fuller description of the advantages of such approach].

In summary, the present study was aimed at investigating the developmental trajectories of word recognition in bilingualism, and how (if at all) language non-selectivity shapes them. Using an adaptation of [@mani2010infant], we presented 20- to 32-months-old toddlers with a cross-linguistic phonological priming paradigm. We manipulated both the phonological overlap between the prime word and target word, and the cognate status of the prime word, to test the hypothesis that bilingual toddlers implicitly name familiar pictures in both languages. If so, phonological priming effect should be larger when the prime label in both languages shares phonological overlap with the target words (*Related/Cognate* condition), compared to when only the label of the prime in one language overlaps phonologically with the target (*Related/Non-cognate*). We also also included a control *Unrelated* condition in which regardless of the cognate status of the prime word, neither of the prime labels in any language overlapped phonologically with the target word.

To test cross-language priming effects, we registered participants' time course of target picture looking in the three conditions. We predicted within-language priming effects, revealed as differences in target looking between *Unrelated* and *Related/Non-cognate* trials, as previously reported [@von2012, @MANI2001, and @MANI2011]. We further predicted cross-language priming effects, revealed as differences in target looking between *Related/Non-cognate* and *Related/Cognate* trials. 
This effect should be stronger in bilinguals, if not completely absent in monolinguals, who should be indifferent to the phonological properties of the stimuli in the other language. 
Following previous studies associating the emergence of phonological links between lexical items to the size of children's vocabulary [@MANI2011], we also compared the explanatory power of two indices of receptive vocabulary size (dominant-language vocabulary size and total vocabulary size) against participants' age. We used parental reports of vocabulary checklist to estimate participants' vocabulary size. We computed vocabulary size in the dominant language (L1 vocabulary) as the proportion of words in the dominant language reported by their caregivers to be understood by the child in a vocabulary checklist, and the total vocabulary size as the proportion of words reported as understood by the child in *both* languages. If within-language or cross-language phonological associations in the lexicon emerge as a function of vocabulary growth, the size of the differences between conditions should larger in participants with larger vocabulary sizes. A model including participants' L1 or total vocabulary size as a predictor should then show better performance than a model that instead includes participants age. In the case of bilinguals, if the emergence of cross-language priming effects is associated with the growth of their vocabulary is *both* languages, a model including total vocabulary size as a predictor should perform better than a model that includes L1 vocabulary size. The opposite pattern would indicate that cross-language priming effects emerge as a function of the growth of L1 vocabulary.

@@

```{r results}
var_labels <- c("b_Intercept" = "Intercept",
				"b_age_std" = "Age (+1 SD)",
				"b_condition1" = "Condition (U vs. RN)",
				"b_condition1" = "Condition (RN vs. RC)",
				"b_lp1" = "Group",
				"sd_child_id__Intercept" = "Intercept",
				"sd_child_id__age_std" = "Age",
				"sd_child_id__lpBilingual" = "Group",
				"sd_child_id:session_id__Intercept" = "Intercept",
				"sd_child_id:session_id__age_std" = "Age",
				"sd_child_id:session_id__lpBilingual" = "Group")

@@

```{r tbl-loos}
loos_tbl[1:5] |> 
	mutate(model = gsub("fit_", "", model)) |>
	gt() |> 
	fmt_number(is.numeric) |> 
	# fmt_markdown(is.character, ) |> 
	cols_label(model = "Model",
			   elpd_diff = "ELPD (diff.)",
			   se_diff = "SE ELPD (diff)",
			   elpd_loo = "ELPD",
			   se_elpd_loo = "SE ELPD")
```

@@

Overall, participants' looking time exceeded chance levels, as indicated by the fact that the 95% HDI of the intercept term excluded zero ($\beta$ = `r coefs$fit_4$b_intercept$mdn_label`, 95% HDI = `r coefs$fit_4$b_intercept$ci_label`). The variability of participants' target preference varied significantly more between participants ($\sigma$ = `r coefs$fit_4$sd_child_id_intercept$mdn_label`, 95% HDI = `r coefs$fit_4$sd_child_id_intercept$ci_label`) than within testing sessions from the same participant ($\sigma$ = `r coefs$fit_4$sd_child_id_session_id_intercept$mdn_label`, 95% HDI = `r coefs$fit_4$sd_child_id_session_id_intercept$ci_label`). The 95% HDI of the coefficient of *Age* had a positive sign, but did not exclude zero ($\beta$ = `r coefs$fit_4$b_age_std$mdn_label`, 95% HDI = `r coefs$fit_4$b_age_std$ci_label`), indicating that participants from all ages showed equivalent overall target word recognition.

The 95% HDI of the coefficient of *Group* in $\mathcal{M}_{1}$ included zero ($\beta$ = `r coefs$fit_4$b_lp1$mdn_label`, 95% HDI = `r coefs$fit_4$b_lp1$ci_label`), indicating an equivalent overall target preference in monolinguals and bilinguals. The 95% HDI of the first contrast of this predictor, comparing *Unrelated* and *Related/Non-cognate* trials, included zero ($\beta$ = `r coefs$fit_4$b_condition1$mdn_label`, 95% HDI = `r coefs$fit_4$b_condition1$ci_label`). The 95% HDI of the second contrast, comparing *Related/Non-cognate* and *Related/Cognate* trials, also included zero ($\beta$ = `r coefs$fit_4$b_condition2$mdn_label`, 95% HDI = `r coefs$fit_4$b_condition2$ci_label`). The overall target preference was equivalent across both pairwise condition comparisons. The interaction term between the first *Condition* contrast contained zero ($\beta$ = `r coefs$fit_4$b_condition1_lp1$mdn_label`, 95% HDI = `r coefs$fit_4$b_condition1_lp1$ci_label`). The interaction term between the second *Condition* contrast also contained zero ($\beta$ = `r coefs$fit_4$b_condition2_lp1$mdn_label`, 95% HDI = `r coefs$fit_4$b_condition2_lp1$ci_label`). The outcomes of this model indicate that lack of differences in overall target looking time across conditions remained across monolinguals and bilinguals.

@@

In Study 1, we adapted an implicit naming paradigm to investigate the potential role of within- and cross-language phonological priming in 21- to 32-month-old children. Overall, participants' performance was equivalent across all priming conditions. Both the overall target looking and its course time across the trial suggested strong word recognition, regardless of participants language profile and age. An exploratory analyses on word recognition times suggested a complex pattern of results, showing an age effect---in which older participants looked at the target above chance earlier than younger infants---and a weak interaction between priming conditions, age, and language group. These results conflict with previous findings in several ways.

First, investigations using implicit naming paradigm had provided strong evidence in favour of infants and toddlers generating internal labels when presented with pictures of familiar objects. @mani2010infant and @mani2011phonological reported priming effects in 18 and 24-month-old infants. In primed trials, participants' target looking preference was influenced by the previous presentation of a phonologically related referent in silence. In control trials---in which prime and target labels were unrelated---no priming effect was found. This suggests that infants implicitly named the prime label, and that the resulting internal phonological form was detailed enough to interact with the subsequent auditory recognition of the target word. In the present study, we did not find evidence in favour of differences in target looking preference between related and unrelated trials, regardless of the age of participants or their language profile.

```{r doe-exposure}
doe_mon_exp <- participants_bcn |>
	filter(lp=="Monolingual")|> 
	unnest_wider(doe) |> 
	mutate(doe2 = ifelse(test_language=="Catalan",
						 doe_spanish, doe_catalan),
		  is_positive =  doe2 > 0) |> 
	count(is_positive) 
```
Both @mani2010infant and @mani2011phonological, and the present study shared a similar design, but sampled different children populations. Participants in Mani and Plunkett's studies were raised in English speaking households in Oxford, while participants in the present study were learning Catalan or Spanish, if not a combination of both. We identify two ways in which this might have influenced the results. One is that the stimuli lists in Catalan and Spanish might have differed in some critical dimension, which interfered infants' ability to generate implicit labels for prime pictures. For example, words in the present study were predominantly bi-syllabic, while those from Mani and Plunkett's studies were mostly monosyllabic, and contained less phonemes on average. Longer word-forms might have been more difficult to retrieve in time for a prime effect to take place in target recognition. Another possibility is that exposure to a second language might have delayed the developmental trajectories of word production in bilinguals, in such way implicit naming emerges later ages. Given that `r scales::percent(doe_mon_exp$n[1] / doe_mon_exp$n[2], accuracy = 0.01)`
of monolinguals in Barcelona were also exposed to a second language, they might as well been affected by such delay in implicit naming trajectories.

The results from the Study 1 also conflict with previous reports of cross-language priming in toddlers. Particularly, @von2012language and @floccia2020translation had provided evidence of phonological priming through translation. One critical different between both studies and the present one is the silent or auditory presentation of the prime labels. In priming through translation studies, the prime is auditorily presented to participants, and cross-language priming effects take place through the subsequent activation of the prime translation. In the present study, the prime label was not named, and participants were expected to generate both the prime label and its translation implicitly. It may be the case that infants in the present study faced a more challenging task, having to internally retrieve phonological forms without any auditory prompt like the one provided by priming-through-translation studies.

Finally, another possibility is that participants in the present study did in fact general prime labels, but such labels did not interfere over target recognition. This could be because they were not phonologically specified enough. This is unlikely, as previous studies have shown strong evidence of fine phonological representations in the early lexicon REFERENCE. Instead, it could be the case that the prime labels were activated, but phonological links between lexical representations are still weak or not present, so that priming effects are not possible until later in age. To our knowledge, this study is the first one to test bilingual children at this age on implicit naming paradigms, so evidence of priming is not present at the moment.

We found no differences between models including participants' age or any of the two indices of vocabulary size (L1 and total vocabulary size) as predictors. Previous studies suggest that same-language vocabulary size predicts better bilingual infants' performance in word recognition tasks, as compared to cross-language vocabulary size [@marchman2010vocabulary].

In summary, the possibly increased difficulty of the task induced by the properties of the stimuli, together with differences in task design and populations might have contributed to the null pattern of results shown in Study 1. Given the complexity of the picture drawn by these factors, we conducted a second study with the aim of replicating the original findings by @mani2010infant and @mani2011phonological. If phonological priming effects are observed in a same-aged population using the adaptation of the task from Study 1, this would provide evidence in favour of cross-language differences in the developmental trajectories of implicit naming. If, on the contrary, a similar patter of null results is found, this would suggest either the existence of finer differences in task implementation compared to the original studies, or point to the effect of interest being less robust than suggested by the literature.

@@

In this study, we conducted a conceptual replication of @mani2010infant and @mani2011phonological. Using the same task from Study 1, we tested a same-aged group of monolinguals being exclusively raised in British English speaking homes, the same population tested in the original studies. As in Study 1, we manipulated the phonological relationship between the prime and target labels at word onset, as well as the cognate status of the prime label. Evidence from implicit naming of prime images would be revealed by differences in target looking or its time course across the trials between *Unrelated* trials, and *Related/Non-cognate* and *Related/Cognate* trials. Given that participants in this study are monolinguals---and contrary to the monolinguals in Barcelona, have received no exposure to any other language---we would not expect the cognate status of the prime labels to exert any effect on subsequent target recognition.

@@

```{r tbl-loos-oxf}
loos_tbl |> 
	mutate(model = gsub("fit_oxf_", "", model)) |>
	gt() |> 
	fmt_number(is.numeric) |> 
	# fmt_markdown(is.character, ) |> 
	cols_label(model = "Model",
			   elpd_diff = "ELPD (diff.)",
			   se_diff = "SE ELPD (diff)",
			   elpd_loo = "ELPD",
			   se_elpd_loo = "SE ELPD")
```

@tbl-loos-oxf summarises the outputs of the LOO-CV model comparisons. Models $\mathcal{M}_{1}$, $\mathcal{M}_{2}$, and $\mathcal{M}_{3}$ showed substantially better predicted performance than model $\mathcal{M}_{0}$, which indicates that adding the *Condition* predictor improves significantly the fit of the models. All $\mathcal{M}_{1}$, $\mathcal{M}_{2}$, and $\mathcal{M}_{3}$ models showed equivalent performance, suggesting that adding the two-way interaction between *Age* and *Condition* did not improve the the fit of the model. We now report a description of the posterior distribution of the regression coefficients of model $\mathcal{M}_{3}$.

Overall, participants' looking time exceeded chance levels, as indicated by the fact that the 95% HDI of the intercept term excluded zero ($\beta$ = `r coefs$fit_3$b_intercept$mdn_label`, 95% HDI = `r coefs$fit_3$b_intercept$ci_label`). The variability of participants' target preference varied significantly more between participants ($\sigma$ = `r coefs$fit_3$sd_child_id_intercept$mdn_label`, 95% HDI = `r coefs$fit_3$sd_child_id_intercept$ci_label`) than within testing sessions from the same participant ($\sigma$ = `r coefs$fit_3$sd_child_id_session_id_intercept$mdn_label`, 95% HDI = `r coefs$fit_3$sd_child_id_session_id_intercept$ci_label`). The 95% HDI of the coefficient of *Age* had a positive sign, but did not exclude zero ($\beta$ = `r coefs$fit_3$b_age_std$mdn_label`, 95% HDI = `r coefs$fit_0$b_age_std$ci_label`), indicating that participants from all ages showed equivalent overall target word recognition.
